---
title: "Reverse Engineering Project"
author: "Sarah Meador"
date: "3/7/24"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [Thousands of kids are missing from school. Where did they go?](https://projects.apnews.com/features/2023/missing-children/index.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

library(tidyverse)
library(janitor)
library(lubridate)

pop_age <- read_csv("data/population-by-age.csv")|>
  clean_names() 
public_enrollment <- read_csv("data/public-enrollment-by-grade.csv")|>
  clean_names() 
private_enrollment <- read_csv("data/private-enrollment-by-grade.csv")|>
  clean_names()

## Load and Cleaning Data

#In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: "Public school enrollment plunged during the pandemic. Our analysis shows where the students went. In the 21 states where data was available, enrollment fell by about 700,000 students between the 2019-2020 and 2021-2022 school years."
-   **Analysis summary**: I am not sure how to subtract the numbers with code, but when I did subtract the numbers I got (2021-2022 school year enrollment minus 2019-2020 enrollment) I did not get the same number from the story. This might be because it seems like there is data from more than 21 states in my data. I got about 2.4 million.

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock
public_enrollment |>
  filter(enrollment >= 0)|>
  group_by(sch_year) |>
  summarize(pub_enrollment = sum(enrollment))
```

### Sentence 2

-   **Sentence text**: "California alone showed over 150,000 missing students in the data, and New York had nearly 60,000."
-   **Analysis summary**: In my analysis, I came to the conclusion that in my updated version of the data, California shows 541,856 missing students from the 2019-2020 school year to the 2021-2022 school year. New York shows 264,796 missing students from these school years in the updated data.

```{r}
# Put code to reverse engineer sentence here
public_enrollment |>
  filter(state == 'CALIFORNIA') |>
  group_by(sch_year) |>
    summarize(total_enrollment = sum(enrollment)) |>
  summarize(enrollment_change = ((max(total_enrollment)) - (min(total_enrollment))))

# Display results of code below this codeblock

```

```{r}
public_enrollment |>
  filter(state == 'NEW YORK') |>
  group_by(sch_year) |>
    summarize(total_enrollment = sum(enrollment)) |>
  summarize(enrollment_change = ((max(total_enrollment)) - (min(total_enrollment))))
```

### Sentence 3

-   **Sentence text**: "That leaves an unexpected 230,000 students still unaccounted for in the data."
-   **Analysis summary**: When looking at the population by school age, from Fall 2021 to Spring 2020, we see a 238,000 population decrease. This confirms their finding but originally it did not seem that this dataset would be the best to showcase this change of school-aged students who may be unaccounted for. The data does not specifically say all of these students are not enrolled in school, so it remains somewhat unclear if it is best for showing their findings.

```{r}
# Put code to reverse engineer sentence here
pop_age |>
  group_by(est_date) |>
  summarize(school_age_pop = sum(population))

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: "The analysis doesn't include data from 29 states, including Texas and Illinois, or the unknown numbers of ghost students who are technically enrolled but rarely make it to class.
-   **Analysis summary**: In my analysis, I could not find evidence that supported the statement, likely because the data has been updated since the story was reported. When I located all the states who provided data, Texas and Illinois were included in my data findings.

```{r}
# Put code to reverse engineer sentence here
all_states <- pop_age |> 
  group_by(state) |> 
  summarise( count_state=n() ) |> 
  arrange(desc(count_state)) 

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
