---
title: "lab_02"
author: "derek willis"
date: "2023-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? **A1. 26 departments, Atmospheric and Oceanic Science**

```{r}
#my attempt
and_climate_seats_1 <- umd_courses |> filter(str_detect(title, "Climate") & seats >= 1) |> 
  group_by(department) |> 
  summarise(
    and_climate_seats_1 = n()
  ) |>
  arrange(desc(and_climate_seats_1))

nrow(and_climate_seats_1)
```

```{r}
#professor answer
umd_courses |>
  filter(str_detect(title, "Climate") & seats >= 1) |>
  group_by(department) |>
  summarize(classes = n()) |>
  arrange(desc(classes))
```

**Q2.** Of the departments listed in A2, which one is the most interesting to you, and why? **A2. It's interesting that there are classes about climate in the Germanic Studies department.**

**Q3** Which departments have the most courses in the Fall 2023 term? And which departments in that term have at least one syllabus posted? How are the top five departments different in each result?

**A3. Departments with most courses (top 5): Business and Management, School of Music, English, Teaching and Learning, Information Studies**

**Departments with syllabus (top 5): Business and Management, Economics, Biological Sciences Program, Information Studies, School of Music**

**Some of the top departments for how many classes offered are not also the same for the ones with the most classes with a syllabus.**

```{r}
#my attempt
umd_courses |>
  filter(term == 202308 & syllabus_count >= 1) |>
  group_by(department) |>
  summarize(classes = n()) |>
  arrange(desc(classes))
```

```{r}
#professor answer
course_count <- umd_courses |>
  filter(term == 202308) |>
  group_by(department) |>
  summarize(course = n()) |>
  arrange(desc(course))

course_count_syllabi <- umd_courses |>
  filter(term -- 202308 & syllabus_count > 0) |>
  group_by(department) |>
  summarize(courses_with_syllabi = n()) |>
  arrange(desc(courses_with_syllabi ))
  
```

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr, saving the result as the same dataframe. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time. **A4. It appears that the spending increases over time.**

```{r}
maryland_expenses |>
  mutate(expenditure_date_new = mdy(expenditure_date)) |>
  mutate(month = floor_date(expenditure_date_new, "month")) |>
  filter(committee_name == 'Moore  Wes For Maryland') |>
  group_by(month) |>
  summarize(total_amount = sum(amount))
```

**Q5.** Using case_when(), create a column in the Maryland expenses data called `spending_location` indicating whether each record spent money in Maryland or outside Maryland. For records that are in Maryland, make the new column's value "In-state" and for the others, make it "Out of state". Then write code that shows the total amount of money spent in each category and describe the results. You can do this in one statement or break it up into two statements (saving the new column to your dataframe). **A5**

```{r}
spending_location <- maryland_expenses |>
  mutate(
    spending_location = case_when(
        str_detect(address, " Maryland ") ~ "In-state",
        .default = "Out of state"
      )
  ) |>
 mutate(
    total = group_by(spending_location) |>
      summarize(total_amount = sum(amount,na.rm=T))
  )
```

**Q6.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab. **A6. Question: How many journalism classes in fall 2023 mention the word "sports" in the class title? Answer: 9**

```{r}
journalism_sports <- umd_courses |> filter(str_detect(title, "Sports") & term == 202308) |> 
  group_by(department) |> 
  summarise(
    journalism_sports = n()
  ) |>
  arrange(desc(journalism_sports))
a
head(journalism_sports)
```
