---
title: "lab_07"
author: "derek willis"
date: "2023-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key if you haven't already. You won't need to install or overwrite it.

```{r echo=FALSE}
#census_api_key("", install=TRUE, overwrite = TRUE)
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2021 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2021 5-year ACS (which is called "acs5") and save them to a dataframe

```{r}
v21 <- load_variables(2021, "acs5", cache=TRUE)
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2021 5-year estimates? You will need to find the variable for median age first (search for it in the 'concept' column at the data frame we just created). **A1 The Maryland jurisdiction with the lowest median age according to the ACS 2021 5-year estimates is Baltimore city with an estimated median age of 35.7.**

```{r}
md_median_age <- get_acs(geography = "county",
              variables = c(medage = "B01002_001"),
              state = "MD",
              year = 2021) |>
arrange(estimate)

```

**Q2** Which Maryland counties have a median age of at least 50? You can use the dataframe you just made to answer this, but you must write code to display only those counties with a median age of at least 50. **A2: Worcester County (50.4) and Talbot County (51.0) both have median age estimates of at least 50.**

```{r}
md_median_age |>
  filter(estimate >= 50)
```

**Q3** We're interested in knowing more about Montgomery County zip codes where overdose calls are coming from and the demographic characteristics they possess. In particular, we want to calculate a rate of calls per 1,000 population. To do this, you'll need to:

1.  Load the Montgomery County 911 overdose calls data.
2.  Create a new dataframe with the total number of calls for each zip code.
3.  Find the ACS variable pertaining to the total 18 and up population (there are several choices here; you want the simplest one).
4.  Get from the ACS data from all Maryland zip codes using that variable.
5.  Join the ACS data to the Montgomery County zip code totals dataframe and rename the estimate column to a meaningful name you've given the ACS variable.
6.  Add a column to the joined dataframe that calculates a rate of calls per 1,000 population.

Which zip code has the highest rate of calls? Which has the lowest?

**A3: Zip code 20877 has the highest rate of calls with about 1.7 calls per thousand people. Zip code 20707 has the lowest rate of calls with about .03 calls per thousand people.**

```{r}
montgomery_overdose_calls <- read_csv("data/montgomery_2022_overdoses.csv") 

montgomery_zip_overdose <- montgomery_overdose_calls |>
  mutate(zip = as.character(zip)) |>
  group_by(zip) |>
  summarize(total_calls = n()) |>
  arrange(desc(total_calls))
  
md_zcta_pop <- get_acs(geography = "zcta",
              variables = c(medage = "B09021_001"),
              state = "MD",
              year = 2019)

moco_pop_overdose <- montgomery_zip_overdose |> inner_join(md_zcta_pop, by=c('zip' = 'GEOID')) |>
  rename(population = estimate)

moco_pop_overdose <- moco_pop_overdose |> 
  mutate(rate_per_1k = (total_calls/population)*1000) |>
  arrange(desc(rate_per_1k))
```

**Q4** Using [CensusReporter.org](https://censusreporter.org/), examine the zip codes with the highest and lowest rates and compare and contrast their demographic profiles. Is there a story here?

**A4: Zip code 20877 has a large Hispanic population, with 44% of residents being Hispanic. Zip code 20707 has a large black population, with 46% of residents being black. Zip code 20707 also has a smaller poverty rate than zip code 20877, with the former having a rate of 8.9% and the latter having a rate of 13.2%. A story could be found here in seeing if there is a trend between poverty levels and overdose calls and exploring why.**

**Q5** Go back to your joined dataframe from Q3 and add two more columns representing the minimum and maximum populations based on the estimate and margin of error, then calculate per capita rates for each. Do you get the same two zip codes for the highest and lowest as you did before? Does that give you more or less confidence in your answers to Q3?

**A5: I got zip code 20707 again for the zip code with the lowest per capita rate of calls and zip code 20877 again for the zip code with the highest per capita rate of calls. This gives me more confidence in my answers to question 3 because I have accounted for the margin of error and still got the same answers.**

```{r}
moco_pop_overdose |>
  mutate(min = (population - moe)) |>
  mutate(max = (population + moe)) |>
  mutate(max_rate_per_1k = (total_calls/min)*1000) |>
  mutate(min_rate_per_1k = (total_calls/max)*1000) |>
  arrange(desc(max_rate_per_1k))
  
```
